---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Lily Botha"
date: "02/17/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(readr)
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- read_csv("https://github.com/DataScience4Psych/DataScience4Psych/raw/main/data/raw-data/dennys.csv")
laquinta <- read_csv("https://github.com/DataScience4Psych/DataScience4Psych/raw/main/data/raw-data/laquinta.csv")
```

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 2

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

...

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

Add exercise headings as needed.
